### 2015-11-02 p00-ch00-intro/intro.tex
Introduction polishing

One of the fundamental elements of actor model~\cite{actors:agha,agha97} is \emph{asynchronous} message passing.
In this approach, interactions between objects are modelled as \emph{non-blocking} messages.
One object, the sender, communicates a message to the other object, the receiver.
At the receiver side, a message eventually leads to a method invocation (in an object-oriented paradigm) or a functional call (in a functional paradigm).
In actor model, the locality of the objects is transparent from the messages.
A system is composed of objects with communicating messages.

@@

With the rise of computation power, a new aspect added to create a triangle: object-oriented, actor model, and \emph{concurrency}.
Concurrency makes it harder to verify programs in terms of correctness of runtime behavior. 
Scheduling of object messages in threads of executions is one of main challenges.
In addition, in the presence of scheduling and thread interleaving, it is important that objects are able to protect their state from concurrent modifications for correctness properties.
In this research, we take advantage of a modelling language that aims to mix the three aspects in addition to the capabilities of formal semantics and verification.

@@

We target to deliver ABS concurrency layer in form of a Java library API.
However, Java language specification (JLS) clearly concludes that Java language is not comptabile and ready to be extended as a functional language with the support of algebraic data types.
This has been the focus of research and development to extend Java to a functional language~\cite{odersky1997pizza,henkel2003discovering,nystrom2003polyglot,bracha1998making}.
Scala language~\cite{odersky2004scala} is a result of the research that is a dynamic, functional, and object-oriented language on JVM.
Although such extensions over Java towards a functional language exist, they lack formal semantics in the context of concurrency and actor model.

We primarily focus on ABS concurrency layer as a Java library API.
We argue that the advantage of a correct semantic translation from ABS concurrency layer to its Java equivalent constructs overcomes the additional support of functional layer but with \emph{partial} formal semantics.
Our approach enables programmers to model their Java systems in terms of ABS concurrency semantics and then create a correct mapping and implementation from ABS to Java using an API.
In addition, we aim to facilitate ways to actually program in ABS models as if the programmer is able to take advantage of standard Java libraries.

In the next section, we sketch a high-level overview of the architecture design goals and how this research plans its objectives.

@@

@@

\paragraph{Scalability} 
The distribution level that asynchronous message passing mechinism deliver
challenges performance efficiency.
In such a context, the number of messsages delivered among actors in the 
environment is not predictable at runtime.
The goal is to ensure the actor system scales in performance with least 
influence from the number of asynchronous messages delivered in the system.
We direct benchmarks on the system to present the performance of the system
using different number of messages delivered in the system.

\paragraph{Modularity} 
The scope of the research spans to a number of layers revolving around ABS
 language:
\begin{itemize}
\item \emph{Compiling ABS to a target programming language}
One first objective it compile an ABS model to a target programmig language.
The compilation can be at \emph{source-to-source} or \emph{source-to-bytecode}
level.
Target languages potentially include mainstream programming languages such as
Java, Erlang, Haskell, and Scala.
Thus, the ABS programmer should be able to use an ABS compiler to compile
their ABS models to the target languages.
\item \emph{Using ABS semantics as an API in a programming language}
ABS language is precisely expressed with structural operational semantics~\cite{johnsen2012abs} in addition to its syntax definition.
Therefore, the semantics of ABS can be delivered in a programming language
as Application Programming Interface (API) as long as the programming language provides sufficient constructs to respect ABS language semantics.
In addition, such ABS programming API should be \emph{verifiable}.
If such ABS mapping to a programming API is provided, a programmer is able to
take advantage of ABS semantics without directory programming in ABS.
Such capability from ABS enables industry users of mainstraim languages to
model their systems in ABS semantics using their programming languages and
platforms.
\item \emph{Modelling in ABS}
ABS language provides solid structure and semantics to model concurrent
and distributed systems.
The user (that can be a programmer, an analysist, or a researcher) rightfully
demands to have access to a tool-set and IDE that allows working with ABS 
models in a user-friendly way.
The ABS IDE and tool-set developers should be able to easily reuse and compose
over existing modules and components.
\end{itemize}

Based on the scopes and layers of usage of this research, we aim to provide 
a highly modular architecture in design and implementation.
The goal is verifiable based on produced artifacts of this research in terms
of their \emph{usability}, \emph{re-usability}, and \emph{composability} in
different contexts.

\paragraph{Extensibility}
Delivering ABS language features as an API brings about a new requirement to be able to configure and extende the API. 
It is a common practice in software engineering that components are developed with a mixture of \emph{configuration}, \emph{extension} and \emph{composition}.
Therefore, the ABS programmer should be able to extend such API to custom requirements. 
The extensibility is provided either through configuration of the API or API extension.

Figure~\ref{fig:arch} presents the general architecture of the implementation.

@@

\subsection{Programming Languages Overview}
\label{sec:intro:proglangs}

@@

In the previous section, we presented an overview of the programming languages and their categories that aim to provide actor model~\cite{agha86book}.
Since many programming languages faced different challenges to provide the necessary syntax and semantics for this at the level of the language, many libraries and frameworks have been born with a motivation to fill this gap.
We observe that the more the language itself is close to the actor model semantics, the less external libraries and frameworks target this gap. 
In the following, we briefly provide a few frameworks
\footnote{A more comprehensive list can be obtained at \cite{KarmaniSA09} and  \url{https://en.wikipedia.org/wiki/Actor_model\#Programming_with_Actors}}
and libraries for Java as one of the main focus areas of this research is to study this gap for a mainstream language such as Java.

@@

One of the main techniques used in libraries to deliver actor programming in Java is byte-code engineering~\cite{dahm1999byte,bruneton2002asm,asm}.
Byte-code engineering modifies the generated byte-code for compiled classes in Java either during compilation or at runtime.
Although, this technique is commonly used and argued to provide better performance optimization~\cite{vallee1999soot}, it introduces deep challenges regarding the verification of the running byte-code~\cite{leroy2001java,leroy2003java}.
Thus, one of the main objectives in this research to provide a \emph{statically} verifiable technique to provide actor programming in Java.

\section{Contributions}
\label{sec:intro:contribs}

Based on what we discuss in objectives and related work, the contributions of this thesis can be summarized as the following table:

\begin{table}[h]
\centering
\begin{tabular}{p{7cm}p{3cm}}
\textsfb{Topic} & \textsfb{Chapter/Section}
\\ \toprule
{Formalization of the mapping from ABS to Java including the operational semantics and ABS co-operative scheduling in Java} & Chapter~\ref{ch:p01:ch01} and \ref{ch:p01:ch02}
\\ \midrule
Design and implementation of ABS concurrency layer in Java & Chapter~\ref{ch:p02:ch01} and \ref{ch:p02:ch02}
\\ \midrule 
Design and implementation of a massive-scale monitoring system based on ABS API in Java & Chapter~\ref{ch:p03:ch01}
\\ \bottomrule 
\end{tabular}
\caption{Thesis Contributions Summary}
\label{tbl:thesis}
\end{table}

>>>>

One of the fundamental elements of actor model~\cite{actors:agha,agha97} is \emph{asynchronous} message passing.
In this approach, interactions between objects are modelled as \emph{non-blocking} messages.
One object, the sender, communicates a message to the other object, the receiver.
At the receiver side, a message eventually leads to a method invocation (in  object-oriented paradigm) or a function call (in functional paradigm).
In actor model, the locality of the objects is transparent from the messages.
A system is composed of objects with communicating messages.

@@

With rapid increase of computation power, a new aspect added to create a triangle: object-oriented, actor model, and \emph{concurrency}.
Concurrency makes it harder to verify programs in terms of correctness of runtime behavior. 
Scheduling of object messages in threads of executions is one of main challenges.
In addition, in the presence of scheduling and thread interleaving, it is important that objects are able to protect their state from concurrent modifications for correctness properties.
In this research, we take advantage of a modelling language that aims to mix the three aspects in addition to the capabilities of formal semantics and verification.

@@

% We deliver ABS concurrency layer in form of a Java library API.
% However, 
Java language specification (JLS)~\cite{gosling2000java} clearly concludes that Java language is not comptabile and ready to be extended as a functional language with the support of algebraic data types.
This has been the focus of research and development to extend Java to a functional language~\cite{odersky1997pizza,henkel2003discovering,nystrom2003polyglot,bracha1998making}.
Scala language~\cite{odersky2004scala} is a result of the research that is a dynamic, functional, and object-oriented language on JVM.
Although such extensions over Java towards a functional language exist, they lack formal semantics in the context of concurrency and actor model.

We primarily focus on ABS concurrency layer and deliver its semantics as a Java API.
We argue that the advantage of a correct semantic translation from ABS concurrency layer to its Java equivalent constructs overcomes the additional support of functional layer with \emph{partial} formal semantics.
Our approach enables programmers to model their Java systems in terms of ABS concurrency semantics and then create a correct mapping and implementation from ABS to Java using an API.
In addition, we aim to facilitate ways to actually program in ABS models as if the programmer is able to take advantage of standard Java libraries.

@@

In this section, we present a high-level overview of design goals that we pursue in this research. 
Equivalently, the effectivness and success of this research is measured upon the achievement of the discussed goals and objectives.

@@

\paragraph{Scalability} 
Asynchronous message passing in ABS is a fit for distributed systems.
In such systems, the number of messsages delivered among actors in the 
environment is not predictable at runtime.
Therefore, distribution challenges performance efficiency or scalability.
The goal is to ensure the actor system scales in performance with least 
influence from the number of asynchronous messages delivered in the system.

\paragraph{Modularity} 
The scope of the research spans to a number of layers revolving around ABS
 language:
\begin{itemize}
\item \emph{Compiling ABS to a target programming language}
One first objective is to compile an ABS model to a target programmig language.
The compilation can be at \emph{source-to-source} or \emph{source-to-bytecode}
level.
Target languages potentially include mainstream programming languages such as
Java, Erlang, Haskell, and Scala.
Thus, the ABS programmer should be able to use an ABS compiler to compile
their ABS models to the target languages.
\item \emph{Using ABS semantics as an API in a programming language}
ABS language is precisely expressed with structural operational semantics~\cite{johnsen2012abs} in addition to its syntax definition.
Therefore, the semantics of ABS can be delivered in a programming language
as Application Programming Interface (API) as long as the programming language provides sufficient constructs to respect ABS language semantics.
In addition, such ABS programming API should be \emph{verifiable}.
If such ABS mapping to a programming API is provided, a programmer is able to
take advantage of ABS semantics without directly programming in ABS.
Such capability from ABS enables industry users of mainstraim languages to
model their systems in ABS semantics using their programming languages and
platforms.
\item \emph{Modelling in ABS}
ABS language provides solid structure and semantics to model concurrent
and distributed systems.
The user (that can be a programmer, an analysist, or a researcher) rightfully
demands to have access to a tool-set and IDE that allows working with ABS 
models in a user-friendly way.
The ABS IDE and tool-set developers should be able to easily reuse and compose
over existing modules and components.
\end{itemize}

% Based on the scopes and layers of usage of this research, we aim to provide 
% a highly modular architecture in design and implementation.
% The goal is verifiable based on produced artifacts of this research in terms
% of their \emph{usability}, \emph{re-usability}, and \emph{composability} in
% different contexts.

\paragraph{Extensibility}
Delivering ABS language features as an API brings about a new requirement: to configure or extend the API. 
It is a common practice in software engineering that components are developed with a mixture of \emph{configuration}, \emph{extension} and \emph{composition}.
Therefore, the ABS programmer should be able to extend such API to custom requirements. 
The extensibility is provided either through configuration of the API or API extension.

\Behrooz{It's worth to include the figure but needs more work to justify its presence and how it relates to this thesis.}
Figure~\ref{fig:arch} presents the general architecture of the implementation.

@@

We discuss a brief overview of related work in the context of programming languages, actor model, and concurrency.
We devide the overview in two levels; one is at the level of the programming languages and the other is for the external (third-party) libraries developed for programming languages.

\subsection{Programming Languages}
\label{sec:intro:proglangs}

@@

% In the previous section, we presented an overview of the programming languages and their categories that aim to provide actor model~\cite{agha86book}.
Since many programming languages faced different challenges to provide the necessary syntax and semantics for actor model and concurrency at the level of the language, many libraries and frameworks have been born with a motivation to fill this gap.
We observe that the more the language itself is close to the actor model semantics, the less external libraries and frameworks target this gap. 
In the following, we briefly provide a few frameworks
\footnote{A more comprehensive list can be obtained at \cite{KarmaniSA09} and  \url{https://en.wikipedia.org/wiki/Actor_model\#Programming_with_Actors}}
and libraries for Java as one of the main focus areas of this research is to study this gap for a mainstream language such as Java.

@@

One of the main techniques used in libraries to deliver actor programming in Java is byte-code engineering~\cite{dahm1999byte,bruneton2002asm,asm}.
Byte-code engineering modifies the generated byte-code for compiled classes in Java either during compilation or at runtime.
Although, this technique is commonly used and argued to provide better performance optimization~\cite{vallee1999soot}, it introduces deep challenges regarding the verification of the running byte-code~\cite{leroy2001java,leroy2003java}.

\section{Oraganization and Contributions}
\label{sec:intro:contribs}

In this thesis, we focus on the concurrency layer of ABS.
We formalize a subset of Java language in regards with ABS.
We provide the operaional semantics of the concurrency layer for scheduling of actors with an extension for deadlines and priorities. 
We deliver ABS concurrency as a Java API.
We utilize Java~8 and Java Concurrency API to preserve the semantics of ABS concurrency layer.
We sketch a general architecture on how ABS language can be modularly translated into different programming languages.
We present how ABS API library can be used in the translation of ABS models into Java.

Table~\ref{tbl:thesis} summarizes the structure of the research and contributions in the chapters of this text:

\begin{table}[h]
\centering
\begin{tabular}{p{7cm}p{3cm}p{3cm}}
\textsfb{Topic} & \textsfb{Part} & \textsfb{Chapter/Section}
\\ \toprule
{Formalization of the mapping from ABS to Java including the operational semantics and ABS co-operative scheduling in Java} & Programming Model (Part~\ref{p:model}) & Chapter~\ref{ch:p01:ch01} and \ref{ch:p01:ch02}
\\ \midrule
Design and implementation of ABS concurrency layer in Java & Implementation (Part~\ref{p:impl}) & Chapter~\ref{ch:p02:ch01} and \ref{ch:p02:ch02}
\\ \midrule 
Monitoring method call sequences using annotations & Application (Part~\ref{p:app}) & Chapter X
\\ \midrule
Design and implementation of a massive-scale monitoring system based on ABS API in Java & Application (Part~\ref{p:app}) & Chapter~\ref{ch:p03:ch01}
\\ \bottomrule 
\end{tabular}
\caption{Thesis Contributions Summary}
\label{tbl:thesis}
\end{table}
